---
title: "P8105 HW2"
author: "Thirsten Stockton"
date: "2022-10-02"
output: github_document
---

# Problem 1

***Importing and Cleaning Data***


LOADING PACKAGES AND SETTING PRINT OPTIONS FOR TIBBLES

```{r, message= FALSE}

library(tidyverse)
library(readxl)
library(lubridate)
options(tibble.print_min = 5)

```

BRINGING IN OUR DATASET

```{r}

nyc_transit_data = 
  read_csv(
    "./NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
  col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c",     Route11 = "c"))%>%
  janitor::clean_names() %>%
  select(
    line,
    station_name,
    station_latitude,
    station_longitude,
    contains("Route"),
    entry,
    vending,
    entrance_type,
    ada)%>%
   mutate(
     entry = ifelse(entry == "YES", TRUE, FALSE))

```

This dataset contains variables pertaining to subway line, station name, station latitude, station longitude, Route#, entry, vending, entrance type and ada compliance. So far, the csv file has been read in, column types where changed for "Route" variables that were coded as numeric, the data set was restricted to the above variables, and entry was changed to a logical variable. The dataframe dimensions are 19x1868. The data is not tidy because the "Route" variables need to be changed from wide format to long format.

***Questions***

How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St 8th Avenue; 125st Broadway; 125st Lenox); the distinct function may be useful here. ***465***

```{r}
nyc_transit_data %>% 
  select(station_name, line) %>% 
    distinct
```

How many stations are ADA compliant? ***84***

```{r}
nyc_transit_data %>% 
  filter(ada == TRUE) %>% 
    select(station_name, line) %>% 
      distinct
```

What proportion of station entrances / exits without vending allow entrance? ***37.7%***

```{r}
nyc_transit_data %>% 
  filter(vending == "NO") %>% 
    pull(entry) %>% 
      mean
```

Reformatted Data: Distinct stations that serve the A train ***60***
```{r}
nyc_transit_data %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line, route_num) %>% 
  distinct
```

Reformatted Data: Distinct stations that serve the A train that are ada compliant ***17***

```{r}
nyc_transit_data %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

# Problem 2

Mr. Trash Wheel

``` {r, message = FALSE}

trash_wheel_df = read_excel("./Trash Wheel Collection Data.xlsx",
    sheet = "Mr. Trash Wheel") %>%
      janitor::clean_names() %>%
        drop_na (dumpster) %>%
           mutate (join = "a") %>%
              mutate(year= as.numeric(year)) %>%
                round(sports_ball, digits = 0)

trash_wheel_df
  
```

Professor Trash Wheel

``` {r}

p_trash_wheel_df = read_excel("./Trash Wheel Collection Data.xlsx",
    sheet = "Professor Trash Wheel") %>%
      janitor::clean_names() %>%
        drop_na (dumpster) %>%
          mutate (join = "b") 
  
        
        
 p_trash_wheel_df     
  

```

***Having issues locating "sports_balls"***

Combining datasets

```{r}

mr_prof_trash_df =
  bind_rows(trash_wheel_df, p_trash_wheel_df)

mr_prof_trash_df

```

Finding total weight collected by Professor Trash Wheel

```{r}

mr_prof_trash_df %>%
  filter( join == "b" ) %>%
    select(weight_tons) %>%
      sum()

```

Finding total number of sports balls collected by Mr. Trash Wheel in 2020

```{r}

mr_prof_trash_df %>%
  filter( join == "a", year == 2020 ) %>%
    select(sports_balls) %>%
      sum()

```


_This dataset has 641 observations and 17 variables. Key varaibales include date of trash collection, and key stats about weight of trash and types of trash collected. The amount of trash collected by Professor Trash Wheel from 2017 to 2022 totals to 190.12 tons. The total amount of sports balls collected by Mr. Trash Wheel in 2020 is 856._

# Problem 3

***Cleaning pols-month data***
```{r}

politician_df = 
  read_csv(
    "./fivethirtyeight_datasets/pols-month.csv") %>%
  separate(mon, sep = "-", into = c("year", "month", "day")) %>%
    mutate( month = month.name[as.numeric(month)]) %>%
      mutate( president = recode(prez_gop, "0 " =  "dem", "1" =           "gop") ) %>%
            select(-prez_gop ) %>%
              select(-prez_dem) %>%
                select(-day) %>%
                  mutate(year= as.numeric(year))


politician_df

```

***Cleaning SNP***

```{r}

snp_df = 
  read_csv(
    "./fivethirtyeight_datasets/snp.csv") %>%
   mutate(date=mdy(date)) %>%
  separate(date, sep = "-", into = c("year", "month", "day")) %>%
    mutate( month = month.name[as.numeric(month)]) %>%
      mutate(year= as.numeric(year))
        
        


snp_df

```

***Tidying Unemployment data***

```{r}

unemp_df = 
  read_csv(
    "./fivethirtyeight_datasets/unemployment.csv") %>%
        pivot_longer(
          Jan:Dec,
          names_to = "month",
          values_to = "Unemployment %" ) %>% 
            rename(year=Year)

unemp_df

  
```

***Final merge***

```{r}

prep_df =
  left_join(politician_df, snp_df, by=c("year", "month"))

prep_df

final_df =
  left_join(prep_df, unemp_df, by=c("year","month"))

final_df

range(final_df$year)

```

This dataset contain information on date, the break down of current American leaders by political party on that given date, what the S&P 500 closed at that day, and what the unemployment rate was. The dimensions are 12x822 and the years covered range from 1947-2015.